---
title: "Machine Learning Project"
author: "John R Ferguson"
date: "January 14, 2019"
output: html_document
---

## Machine Learning Project:  Prediction Assignment 

Background

In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

## Project Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(randomForest)
library(caret)
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

## Data analysis/ preprocessing

```{r data analysis}
# Remove variables with a lot of NAs
maxNACount <- nrow(training) / 2000
removeColumns <- which(colSums(is.na(training) | training=="") > maxNACount)
trainclean <- training[,-removeColumns]
testclean <- testing[,-removeColumns]
# Remove variables with timestamps
removeColumns <- grep("timestamp", names(trainclean))
trainclean2 <- trainclean[,-c(1, removeColumns )]
testclean2 <- testclean[,-c(1, removeColumns )]
# Remove header/index variables
remove=c('user_name','new_window','num_window')
trainclean3 <- trainclean2[, -which(names(trainclean2) %in% remove)]
testclean3 <- testclean2[, -which(names(testclean2) %in% remove)]
```

## Data modeling

```{r data modeling}
# split data into a train set and a test set
inTrain = createDataPartition(trainclean3$classe, p = 0.75, list=FALSE)
MyTrain = trainclean3[inTrain,]
MyTest = trainclean3[-inTrain,]

set.seed(26587)
# run random forest model
modelRF <- randomForest(classe ~ ., data = MyTrain[-1], importance=TRUE,ntrees=10)
modelRF
```

## testing & validation

```{r test & validation}
ptraining <- predict(modelRF,MyTrain)
print(confusionMatrix(ptraining,MyTrain$classe))

pval <- predict(modelRF,MyTest)
print(confusionMatrix(pval,MyTest$classe))

```

## prediction

```{r prediction}
ptest <- predict(modelRF,testclean3)
ptest
```
